A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2021-10-22 paper/elsarticle/elsarticle-template.tex
paper

Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene~\cite{schlund2018moglichkeiten,uva2018evaluating,masood2019augmented,gattullo2019towards,aschenbrenner2019comparing,mayrhofer2019one,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to such instructions in paper form or on a statically situated screen. Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed typically using some form of handheld or head-mounted display. Handheld devices such as tablets do not enable the user to work hands-free while using the device, or force the user to wear additional gear. Reliance on head-mounted displays has two adverse consequences of its own: a head-mounted display must be worn by each individual wishing to partake in the augmentation, and such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive. A HoloLens 2, for instance, is such a head-mounted display; at a weight of 566~grams \cite{hololens}, it places an additional ergonomic load on the worker, in particular if worn over a longer duration or in bent positions.

@@

Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene. Since our approach to projector calibration relies on a calibrated camera, what remains of this section is concerned primarily with the recovery of 2D-3D correspondences first in support of calibrating cameras, then of projectors. A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the local scene plane.

\begin{figure}
    \centering
    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlaid).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
    \qquad
    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlaid).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
    \caption{Recovering 2D positions in support of projector calibration. (a) The asymmetrical circle pattern that we project to each of the target locations in the scene plane. 2D positions of the 2D-3D correspondences to be used for calibrating the projector are detected automatically, as the respective circle centers in the asymmetrical circle pattern image and given in a fixed ordering. (b) For each such target location, a projector calibration image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected using the same algorithm as in (a). A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected. Detected 2D projected asymmetrical circle pattern circle center points and chessboard corners overlain for illustration.}
    \label{fig:2d}
\end{figure}

\paragraph{2D-3D correspondences for camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units in terms of which the camera calibration is carried out, and---owing to how our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed. 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.

\paragraph{2D-3D correspondences for projector calibration} As in camera calibration, we calibrate the projector by relying on 2D-3D correspondences. For projector calibration, however, obtaining those correspondences involves \textit{projecting} a calibration pattern to the ground plane. The 2D pixel positions of the projected pattern are obtained by

 are known \textit{a priori}      that we detect using the calibrated downwards-facing camera.

 The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle center points in the asymmetrical circle pattern in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector. We project the asymmetrical circle pattern to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each. Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}). Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.

\begin{figure}
    \centerline{\includegraphics[scale=.375]{images/2d3d.png}}
    \caption{Obtaining 2D-3D correspondences for projector calibration, using a projector calibration image (cf.\ Figure~\ref{fig:2d}(b)). First, the scene plane (gray) is recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector (red)---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane. The downwards-facing camera is shown in terms of its camera frustum (black, with up vector in red).}
    \label{fig:3d}
\end{figure}

We recover the scene plane via spatial resection by applying a PnP algorithm \cite{collins2014infinitesimal} to the 2D-3D correspondences obtained using a chessboard pattern. Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we account for the possibility of an uneven floor by placing a chessboard pattern in close proximity to the projected asymmetrical circle pattern in each projector calibration image in order to recover the scene plane locally to each target location (cf.\ Figure~\ref{fig:2d}(b)). Note that in principle, we could project a chessboard pattern instead of an asymmetrical circle pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.

The pose of the projector, for each projector calibration image, is provided relative to the 3D points of the pattern---and thus in the coordinate frame of the camera---alongside $\mathtt{K}_\text{proj}$ by the aforementioned optimization procedure. Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location). This is sufficient for our needs in Section~\ref{sec:approach:homography}, and it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.png}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. A physical camera placed to face downwards is almost certain to not face downwards precisely. We correct for this possibility by rotating the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the plane's normal. Since it is in accordance with our downwards-facing camera the we wish to orient our virtual projector and since our virtual projector is required to face directly downwards to the scene plane to compute our corrective image warp, it is the orientiation of the virtual camera that we use to subsequently orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/1_big.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. The virtual projector is then obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

@@

If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion model coefficients) can be placed elsewhere relative to the scene plane. If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp. Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the original image to the scene plane from the viewpoint of the virtual projector. This warp can be effected using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.

\begin{figure}
    \centering
    \subfloat[\centering Projecting original image.]{{\includegraphics[height=4.2cm]{images/proj_projim.png}}}
    \qquad
    \subfloat[\centering Projecting our warped image.]{{\includegraphics[height=4.2cm]{images/proj_projim_waped.png}}}
    \caption{Projection to the scene plane from the recovered projector viewpoint (bottom left, black; virtual projector in center, gray) of the original image and of the warped image. (a) Projecting the original image to the scene plane. (b) After warping the original image according to our plane-induced homography for the given target location, the image is projected in a manner that appears free of projective distortions, aligned with the axes of the virtual camera (via the virtual projector), and to have the desired dimensions in the scene plane, expressed in metric units. Note that unprojected background is shown set to black.}
    \label{fig:warp}
\end{figure}

@@

\paragraph{Virtual projector} The placement of the virtual projector is to determine from which pose the image to be projected is to \textit{appear} to have been projected. We carry out this placement according to a small handful of steps. First, we (i) rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the back-projection of the center of the image to be projected.} with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}). Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane) and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector. Finally, we (iii) align the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ again Figure~\ref{fig:virtualproj}). The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions. We additionally (iv) adjust the height above the scene plane of the virtual projector, in order to satisfy desired projected image dimensions provided in metric units.

Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of the downwards-facing camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed. Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement of augmentations.

\begin{figure}
    \centering
    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.3cm]{images/system2.png}}}
    \quad
    \subfloat[\centering Example projection of image warped for target location using our approach.]{{\includegraphics[height=3.3cm]{images/hans_undist_inset.png}}}
    \caption{Evaluation scenario. (a) Our hardware setup, comprised of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and a Stereolabs Zed 2 stereo camera, of which we used only the left view. (b) Example augmentation produced by projecting---to one of our 15 target locations---an image warped using our approach, as seen from the downwards-facing camera. Note that the warp places the axes of the augmentation in accordance with the axes of the camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm). Corners of the full projected image extent projected to the floor in light green.}
    \label{fig:eval}
\end{figure}

@@

\begin{table}[ht]
\caption{Deviations in measured lengths from 50~cm~$\times$~31.25~cm across the 15 augmentations produced using our approach, provided in cm and in pixel units of the downwards-facing camera (pixel size 3.5 mm). Note that in no instance do mean and standard deviation exceed 1 pixel from the viewpoint of the camera.}
\label{table:length}
\centering
{\small
\begin{tabularx}{7.4cm}{r p{1cm} p{1cm} p{1cm} p{1cm}}
\toprule
  & Top & Bottom & Left & Right \\
\midrule
Mean {\tiny[cm]} & 0.03 & 0.07 & 0.09 & 0.06 \\
Stdev {\tiny[cm]} & 0.33 & 0.17 & 0.17 & 0.17 \\
\midrule
Mean {\tiny[px]} & 0.1 & 0.19 & 0.26 & 0.18 \\
Stdev {\tiny[px]} & 0.95 & 0.48 & 0.49 & 0.49 \\
\bottomrule
\end{tabularx}}
\end{table}

\begin{table}[ht]
\caption{Deviations in measured angles from $90^\circ{}$ across the 15 augmentations produced using our approach.}
\label{table:angle}
\centering
{\small
\begin{tabularx}{6.75cm}{r p{1.95cm} p{1.95cm}}
\toprule
  & Top Left & Bottom Right \\
\midrule
Mean {\tiny[deg]} & -0.03 & 0.0 \\
Stdev {\tiny[deg]} & 0.4 & 0.34 \\
\bottomrule
\end{tabularx}}
\end{table}

>>>>

Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene~\cite{schlund2018moglichkeiten,uva2018evaluating,masood2019augmented,gattullo2019towards,aschenbrenner2019comparing,mayrhofer2019one,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to such instructions in paper form or on a statically situated screen. Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed typically using some form of handheld or head-mounted display. Handheld devices such as tablets do not enable the user to work hands-free while using the device, or force the user to wear additional gear. Reliance on head-mounted displays has two adverse consequences of its own: a head-mounted display must be worn by each individual wishing to partake in the augmentation, and such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive. A HoloLens 2 (Microsoft Corporation, USA), for instance, is such a head-mounted display; at a weight of 566~grams \cite{hololens}, it places an additional ergonomic load on the worker, in particular if worn over a longer duration or in bent positions.

@@

Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; if proceeding accordingly, the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene. What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the downwards-facing camera (used to obtain the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (used for obtaining the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).

\paragraph{2D-3D correspondences for camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern, and is such a pattern that we use as well. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units of the camera calibration, and---since our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed. 2D-3D correspondences are then recovered for each calibration image, respectively, and the resulting list of correspondences is used to recover the camera calibration matrix~$\mathtt{K}_\text{cam}$.

\begin{figure}[ht]
    \centering
    \subfloat[\centering Projector calibration image (detections overlain).]{{\includegraphics[height=3cm]{images/detections.png} }}
    \qquad
    \subfloat[\centering 3D positions for projector calibration (red, in scene plane).]{{\includegraphics[height=3cm]{images/2d3d.png}}}
    \caption{Obtaining the 3D positions of the 2D-3D correspondences for projector calibration. (a) For each target location, a projector calibration image is acquired from the viewpoint of the downwards-facing camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby. Automatically detected circle pattern center points and chessboard corners overlain for illustration. (b) Given a projector calibration image, the scene plane (gray) is recovered in the coordinate frame of the camera (black, up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used for calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown). The downwards-facing camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
    \label{fig:2d}
\end{figure}

\paragraph{2D-3D correspondences for projector calibration} For projector calibration, we obtain the 2D-3D correspondences by \textit{projecting} a calibration pattern to the scene plane, once for each target location. Since we know that the 3D positions of the projected pattern must lie in the scene plane, we can recover the 3D correspondence of a pattern point detected in the 2D image plane of the downwards-facing camera by intersecting the back-projection of the 2D position with the scene plane. We reuse the chessboard pattern---used above for camera calibration---to recover the scene plane locally to each target location via spatial resection \cite{Hartley2004}, by placing the chessboard pattern on the floor near the projected pattern and using the resulting 2D-3D chessboard correspondences as input to a PnP algorithm \cite{collins2014infinitesimal}. While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we recover a scene plane per projector calibration image to account for the possibility of an uneven floor. Since our projector calibration images thus already contain a chessboard, we choose an alternative calibration pattern as the pattern to project, namely an asymmetric circle pattern. We again use a specialized algorithm \cite{bradski2000opencv} to obtain the circle centers of the asymmetric circle pattern, capable of recovering circle centers even under projective distortions.

Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(a)); given the scene plane and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:2d}(b)). We obtain the 2D positions of the 2D-3D correspondences by simply running the algorithm for detecting circle centers on the original asymmetric circles pattern image (i.e., the one we feed to the projector). Since---as in the case of recovering chessboard pattern corners---the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain respective 2D-3D correspondences for each projector calibration image, which we use to recover the projector calibration matrix~$\mathtt{K}_\text{proj}$.

Recall that the pose~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$ of the projector, for each projector calibration image~$j$, is provided alongside $\mathtt{K}_\text{proj}$, relative to the 3D points of the pattern and thus in the coordinate frame of the downwards-facing camera. Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location). This is sufficient for our needs in Section~\ref{sec:approach:homography}, and it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. It is in accordance with the $X$- and $Y$-axes of the downwards-facing camera the we wish to orient our virtual projector. To avoid projective distortions, the virtual projector is required to face directly downwards to the scene plane; a physical camera placed to face downwards, however, is almost certain to not face downwards 	precisely. Accordingly, we rotate the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal with respect to the minimum arc-length rotation. It is then the $X$- and $Y$-axes of the \textit{virtual} camera (gray) that we subsequently use to orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion model coefficients) can be placed elsewhere relative to the scene plane. If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp. Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the original image to the scene plane from the viewpoint of the virtual projector. This warp can be effected using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_proj.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. The virtual projector is then obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

\paragraph{Virtual projector} The placement of the virtual projector determines from which pose the image to be projected is to \textit{appear} to have been projected. We carry out this placement according to a small handful of steps. First, we rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual camera} facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}). It is to the virtual camera's $X$- and $Y$-axes that we subsequently set the $X$- and $Y$-axes of the virtual projector. To obtain the virtual projector, we (i) intersect the scene plane with the optical axis and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera. Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired projected image dimensions provided in metric units. The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane with the desired scale.

It is, strictly speaking, in the sense that we set the $X$- and $Y$-axes of the virtual projector to the $X$- and $Y$-axes of the virtual camera that we place the horizontal and vertical axes of the desired augmentations in accordance with the horizontal and vertical image axes of the camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed. Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector. A steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, thereby facilitating placement of target locations.

\begin{figure}
    \centering
    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projecting the original image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray). (a) Original image. (b) Projecting the original image to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
    \label{fig:proj}
\end{figure}

\begin{figure}
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray). (a) Original image warped using our approach. (b) After warping the original image according to our approach for the given target location, the image is projected with the effect of having the original image from the viewpoint of the virtual projector. This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and with the desired dimensions in the scene plane, expressed in metric units. Note that pixels shown in black will not be projected to the scene.}
    \label{fig:warp}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.3cm]{images/system2.png}}}
    \quad
    \subfloat[\centering Example projection of image warped for target location using our approach.]{{\includegraphics[height=3.3cm]{images/hans_undist_inset.png}}}
    \caption{Evaluation scenario. (a) Our hardware setup, comprised of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and a Stereolabs Zed 2 stereo camera (extreme front, black), of which we used only the left view. (b) Example augmentation produced by projecting---to one of our 15 target locations---an image warped using our approach, as seen from the downwards-facing camera. Note that the warp places the axes of the augmentation in accordance with the axes of the camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm). Corners of the full projected image extent projected to the floor in light green.}
    \label{fig:eval}
\end{figure}

@@

\begin{table}[ht]
\caption{Means and standard deviations (in units of centimeters and pixels, respectively) of deviations in measured lengths from 50 cm for top and bottom and from 21.25 cm for left and right, with respect to the 15 augmentations produced using our approach (intended augmentation dimensions 50~cm~$\times$~31.25~cm). Pixel size given the downward-facing camera's height above the ground plane was 3.5 mm; note that in no instance do mean and standard deviation exceed 1 pixel from the viewpoint of the camera.}
\label{table:length}
\centering
{\small
\begin{tabularx}{7.4cm}{r p{1cm} p{1cm} p{1cm} p{1cm}}
\toprule
  & Top & Bottom & Left & Right \\
\midrule
Mean {\tiny[cm]} & 0.03 & 0.07 & 0.09 & 0.06 \\
Stdev {\tiny[cm]} & 0.33 & 0.17 & 0.17 & 0.17 \\
\midrule
Mean {\tiny[px]} & 0.1 & 0.19 & 0.26 & 0.18 \\
Stdev {\tiny[px]} & 0.95 & 0.48 & 0.49 & 0.49 \\
\bottomrule
\end{tabularx}}
\end{table}

\begin{table}[ht]
\caption{Means and standard deviations of deviations in measured angles from $90^\circ{}$ with respect to the 15 augmentations produced using our approach.}
\label{table:angle}
\centering
{\small
\begin{tabularx}{6.75cm}{r p{1.95cm} p{1.95cm}}
\toprule
  & Top Left & Bottom Right \\
\midrule
Mean {\tiny[deg]} & -0.03 & 0.0 \\
Stdev {\tiny[deg]} & 0.4 & 0.34 \\
\bottomrule
\end{tabularx}}
\end{table}
